Modeling Two-Dimensional Vesicle Hydrodynamics with Hydrophobic Attraction Potential Simulations

### 2021-08-18 doc/main_jfm.tex
updated table
3.3 revision

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Imposed Forces}
The imposed forces and torques contain two parts. The first part comes
from the hydrophobic attraction potential (HAP). This potential was
introduced in~\cite{Fu20} and models hydrophobic attraction by solving
the screened Laplace equation boundary value problem
\begin{alignat}{2}
  \label{eq:SL}
-\rho^2 \Delta u + u &=0,            && \xx \in \Omega,\\
\label{eq:SLbc}
u(\xx) &= f_i(\xx),\qquad  && \xx \in \Gamma_i, \\
\label{eq:SLff}
u &\to 0,                          &&|\xx| \to \infty,
\end{alignat}
where $f_i$ is a material label with $f_i = 0$, respectively $1$, representing
hydrophilic, respectively hydrophobic, portions of the surface, and $\rho$ is the
decay length of attraction. The forces and torques for this attraction
are 
\begin{align}
  \label{eq:hydrophobicAttraction}
  \FF_i^{\text{hydro}} = \int_{\Gamma_i} {\bf T}\cdot \nnu_i \, \dif s, 
    \quad 
  T_i^{\text{hydro}} = \int_{\Gamma_i} (\xx - \aa_i)^{\perp} \cdot ({\bf T} \cdot \nnu_i) \dif s,
\end{align}
where
\begin{align}
  \label{eq:stress}
\mathbf{T}
= \gamma\rho^{-1}u^2 \mathbf{I} + 2\rho\gamma \left(\tfrac{1}{2}|\nabla
  u|^2 \mathbf{I} - \nabla u  \nabla u^T\right)
\end{align}
is the hydrophobic stress tensor. The parameter $\gamma$ is interfacial
tension. The forces $\FF_i^{\text{hydro}}$ and torques
$T_i^{\text{hydro}}$ are responsible for forming particle aggregates
that partially sequester the hydrophobic portions of the particle
surface.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{HAP Boundary Integral Equation}
Similar to our previous work~\cite{Fu20}, we represent the HAP as a
double-layer potential
\begin{align}
  \label{eq:HAP_DLP}
  u(\xx) = \frac{1}{2\pi} \int_{\bd\Omega} \pderiv{}{\nnu_\yy}
    K_0 \left(\frac{|\xx - \yy|}{\rho}\right) \sigma(\yy) \, \dif s_\yy,
    \quad \xx \in \Omega,
\end{align}
where $K_0$ is the zeroth-order modified Bessel function of the first
kind. By requiring that the density function $\sigma$ satisfies the
second-kind integral equation
\begin{equation}
\label{eq:screenedSKIE}
  f(\xx) = \frac{1}{2}\sigma(\xx) + 
    \frac{1}{2\pi}\int_{\bd\Omega} \pderiv{}{\nnu_\yy}
    K_0 \left(\frac{|\xx - \yy|}{\rho}\right) \sigma(\yy) \, \dif s_\yy,
    \quad \xx \in \bd\Omega,
\end{equation}
the HAP double-layer potential~\eqref{eq:HAP_DLP} satisfies the screened
Laplace equation~\eqref{eq:SL}--\eqref{eq:SLff}. We
solve~\eqref{eq:screenedSKIE} with a collocation method. The integrals
are approximated with either the trapezoid rule or with the
interpolation-based near-singular integration scheme. This results in a
$NN_b \times NN_b$ linear system that is solved with a block-diagonal
preconditioned GMRES solver.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Calculating the Hydrophobic Force}
\label{subsec:calculating_force}
Once \eqref{eq:SKIE} has been solved for $\sigma$, we need to evaluate
the integrals~\eqref{eq:hydrophobicAttraction}. These integrals involve
the stress \eqref{eq:stress} which contains a singular integral for the
gradient of the double-layer potential. To avoid singular integrals, we
use the alternative forms
\begin{align}
  \label{eq:recipforcetorque}
  \FF_i^{\text{hydro}} = \int_{\Gamma_i} \JJ_i \,\dif s,\quad
  T_i^{\text{hydro}}    = \int_{\Gamma_i} 
    (\xx - \aa_i)^{\perp} \cdot \JJ_i  \,\dif s,
\end{align}
where
\begin{align}
  \label{eq:jumpstress1}
  \JJ_{i} = 2\gamma\rho^{-1} \sigma v_i \nnu + 
    2\gamma\rho \frac{d\sigma}{ds} \frac{dv_i}{ds} \nnu -
    2\gamma\rho \frac{d\sigma}{ds} \frac{dv_i}{d\nu} \ttau.
\end{align}
Here, $\ttau$ and $\frac{d}{ds}$ are the unit tangent  and arc length
derivative for $\Gamma_i$, respectively, and
\begin{equation}
\label{eq:uidef}
    v_i(\xx) = u(\xx) - \frac{1}{2\pi} \int_{\Gamma_i} \pderiv{}{\nnu_\yy}
    K_0 \left(\frac{|\xx - \yy|}{\rho}\right) \sigma(\yy) \, \dif s_\yy,
    \quad \xx \in \Omega.
\end{equation}
That is, $\nu_i(\xx)$ is the double-layer potential~\eqref{eq:HAP_DLP},
except that $\Gamma_i$ is excluded from $\bd\Omega$. The advantage in
using \eqref{eq:jumpstress1} over \eqref{eq:stress} is that components
$\sigma$, $v_i$, $\frac{d\sigma}{ds}$, and $\frac{dv_i}{ds}$ are all
smooth.

@@

Let $U_i$ be the interior of the particle indexed by $i$. For $\xx_0
\in \Gamma_i$ and an arbitrary function $g(\xx)$, the notation
\begin{align}
  \jump{g}(\xx_0) = \lim_{\substack{\xx \to \xx_0 \\ \xx \in U_i}}g(\xx)  - 
                    \lim_{\substack{\xx \to \xx_0 \\ \xx \in U^c_i}}g(\xx),
\end{align}
denotes the jump of the limits of $g(\xx)$ taken from the outside and
the inside of $\Gamma_i$.
%\begin{align}
%\lim_{\xx \to \xx_0^+}g(\xx),\quad 
%\lim_{\xx \to \xx_0^-}g(\xx), \quad
%  \jump{g}(\xx_0) = \lim_{\xx \to \xx_0^+}g(\xx)  - \lim_{\xx \to \xx_0^-}g(\xx)
%\end{align}
%denotes the limit taken from points in $U_i^c$,  the limit taken from
%points in $U_i$, and the jump, respectively.  
\begin{lemma}
\begin{align}
  \label{eq:prejump}
  \FF_i = \int_{\Gamma_i} \jump{{\bf T}_2  \nu_i}  \, \dif s,\quad
  T_i = \int_{\Gamma_i} (\xx - \aa_i)^{\perp} \cdot \jump{{\bf T}_2 \nu_i} \, \dif s.
\end{align}
\end{lemma}
\begin{proof}
To show~\eqref{eq:prejump}, we expand
\begin{align}
  \FF_i = \int_{\Gamma_i} {\bf T}_1\nnu + 
    {\bf T}_2\nnu + {\bf T}_3\nnu\,\dif s.
\end{align}
By~\eqref{eq:uidef} and~\eqref{eq:decompdivfree}, we have that $u_i$ is
smooth and $\nabla \cdot {\bf T}_1 = \mathbf{0}$ in $\mathbb{R}^n
\setminus U_i$. Similarly, $v_i = u - u_i$ is smooth and $\nabla \cdot
{\bf T}_3 = \mathbf{0}$ in $U_i$. By the divergence theorem,  
\begin{align}
  \int_{\Gamma_i}  {\bf T}_1\nnu \,\dif s
  = -\int_{\mathbb{R}^n \setminus U_i} \nabla \cdot {\bf T}_1 \,\dif \xx = \mathbf{0},\quad
    \int_{\Gamma_p}  {\bf T}_3 \nnu\,\dif s
  = \int_{U_p} \nabla \cdot {\bf T}_3 \,\dif \xx = \mathbf{0}.
\end{align}
Finally, $\nabla \cdot {\bf T}_2= 0$ in $U_i$. This gives
\begin{align}
  \mathbf{0} = \int_{U_i} \nabla \cdot {\bf T}_2 \, \dif \xx = -\int_{\Gamma_i}  
    ({\bf T}_2 \nnu)^-\,\dif s,
\end{align}
where the superscript denotes the limit taken from in $U_i$.  
Combining the previous four equations gives the first equation in \eqref{eq:prejump}.
The argument for the second equation in \eqref{eq:prejump} is identical. 
\end{proof}

\begin{lemma}
\begin{equation}
\label{eq:postjump}
  \jump{{\bf T}_2\nnu} =  \JJ_i.
\end{equation}
\end{lemma}
\begin{proof}
We have the jump relations for \eqref{eq:uidef}:
\begin{align}
\label{eq:DLjump}
\lim_{\xx \to \xx_0^\pm } u_i(\xx) &= \pm\frac{1}{2}\sigma(\xx_0) + u_i (\xx_0),\\
\label{eq:DLdevjump}
\lim_{\xx \to \xx_0^\pm } \nabla u_i(\xx) &= \pm\frac{1}{2}\frac{\dif \sigma}{\dif s}(\xx_0)\ttau(\xx_0) + \nabla u_i(\xx_0).
\end{align}
Using \eqref{eq:DLjump}, \eqref{eq:DLdevjump}, \eqref{eq:Tsplit}, and
that $v_i$ is smooth in a neighborhood of $\Gamma_i$, we have
\begin{align*}
  \jump{{\bf T}_2\nnu}   &= \jump{(\mathbf{S}(u_i,v_i)  +\mathbf{S}(v_i,u_i))\nnu} \\
  &= \jump{( 2\gamma\rho^{-1} u_i v_i \mathbf{I} + 2\gamma\rho \nabla u_i \cdot \nabla v_i \mathbf{I} 
- 2\gamma\rho \nabla u_i  \nabla v_i^T - 2\gamma\rho \nabla v_i \nabla u_i^T)  \nnu}\\
&= 2\gamma\rho^{-1} \sigma v_i \nnu + 2\gamma\rho \frac{d\sigma }{ds}\frac{dv_i }{ds} \nnu
- 2\gamma\rho \ \frac{d\sigma }{ds} \frac{dv_i }{d\nu} \ttau = \JJ_i,
\end{align*}
as required. 
\end{proof}
  Combing  \eqref{eq:prejump}  with \eqref{eq:postjump} completes the proof of \eqref{eq:recipforcetorque}.

@@

\begin{table}
\caption{Friction Coefficients}
\centering    
\begin{tabular}{c c c c c c c c }
 $\chi$ & 0.0020   &  0.0025 &  0.0030 &  0.0035 &  0.0040 & 0.0045 & 0.0050  \\
\hline                    
$\dot\gamma$ (ns$^{-1}$)        & 0.0033 & 0.0041 & 0.0049 & 0.0057 & 0.0066 & 0.0074 & 0.0082\\
%$F_{f}$ (pN nm$^{-1}$)           & 1.84 & 4.90 & 11.36 & 8.16 & 6.14 &  17.69 &  24.33 \\[1ex]
%$\dot\theta$ (nm ns$^{-1}$)           & $6.84\times10^{-5}$ & $1.31\times10^{-4}$ & $3.46\times10^{-4}$ & $2.32\times10^{-4}$ & $1.81\times10^{-4}$ & $4.91\times10^{-4}$ & $5.83\times10^{-4}$\\ [1ex]
$b$ (pN$\cdot$ns nm$^{-2}$)   & 1.11 & 1.56 & 1.37 & 1.47 & 1.43 & 1.51 & 1.76\\ 
\hline    
\end{tabular} 
\label{table1}
\end{table}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Imposed Forces}
The imposed forces and torques contain two parts. The first part comes
from the hydrophobic attraction potential (HAP). This potential was
introduced in~\cite{Fu20} and models hydrophobic attraction by solving
the screened Laplace equation boundary value problem
\begin{alignat}{2}
  \label{eq:SL}
-\rho^2 \Delta u + u &=0,            && \xx \in \Omega,\\
\label{eq:SLbc}
u(\xx) &= f_i(\xx),\qquad  && \xx \in \Gamma_i, \\
\label{eq:SLff}
u &\to 0,                          &&|\xx| \to \infty,
\end{alignat}
where $0 \leq f_i \leq 1$ is a material label with $f_i = 0$, respectively $1$, representing
hydrophilic, respectively hydrophobic, portions of the surface.  We assume that the $f_i$ and $\Gamma_i$ are
smooth.  The parameter $\rho$ is the
decay length of attraction. The forces and torques for this attraction
are 
\begin{align}
  \label{eq:hydrophobicAttraction}
  \FF_i^{\text{hydro}} = \int_{\Gamma_i} {\bf T}\cdot \nnu_i \, \dif s, 
    \quad 
  T_i^{\text{hydro}} = \int_{\Gamma_i} (\xx - \aa_i)^{\perp} \cdot ({\bf T} \cdot \nnu_i) \dif s,
\end{align}
where
\begin{align}
  \label{eq:stress}
\mathbf{T}
= \gamma\rho^{-1}u^2 \mathbf{I} + 2\rho\gamma \left(\tfrac{1}{2}|\nabla
  u|^2 \mathbf{I} - \nabla u  \nabla u^T\right)
\end{align}
is the hydrophobic stress tensor. The parameter $\gamma$ is interfacial
tension. The forces $\FF_i^{\text{hydro}}$ and torques
$T_i^{\text{hydro}}$ are responsible for forming particle aggregates
that partially sequester the hydrophobic portions of the particle
surface.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{HAP Boundary Integral Equation}
Similar to our previous work~\cite{Fu20}, we represent the HAP as a
double-layer potential
\begin{align}
  \label{eq:HAP_DLP}
  u(\xx) = \frac{1}{2\pi} \int_{\bd\Omega} \pderiv{}{\nnu_\yy}
    K_0 \left(\frac{|\xx - \yy|}{\rho}\right) \sigma(\yy) \, \dif s_\yy,
    \quad \xx \in \mathbb{R}^2,
\end{align}
where $K_0$ is the zeroth-order modified Bessel function of the first
kind and the integral is taken in the sense of principle value whenever $\xx \in \partial \Omega$. 
By requiring that the density function $\sigma$ satisfies the
second-kind integral equation
\begin{equation}
\label{eq:screenedSKIE}
  f(\xx) = \frac{1}{2}\sigma(\xx) + 
    \frac{1}{2\pi}\int_{\bd\Omega} \pderiv{}{\nnu_\yy}
    K_0 \left(\frac{|\xx - \yy|}{\rho}\right) \sigma(\yy) \, \dif s_\yy,
    \quad \xx \in \bd\Omega,
\end{equation}
the HAP double-layer potential~\eqref{eq:HAP_DLP} satisfies the screened
Laplace equation~\eqref{eq:SL}--\eqref{eq:SLff}. We
solve~\eqref{eq:screenedSKIE} with a collocation method. The integrals
are approximated with either the trapezoid rule or with the
interpolation-based near-singular integration scheme. This results in a
$NN_b \times NN_b$ linear system that is solved with a block-diagonal
preconditioned GMRES solver.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Calculating the Hydrophobic Force}
\label{subsec:calculating_force}
Once \eqref{eq:SKIE} has been solved for $\sigma$, we need to evaluate
the integrals~\eqref{eq:hydrophobicAttraction}. These integrals involve
the stress \eqref{eq:stress} which contains a singular integral for the
gradient of the double-layer potential. To avoid singular integrals, we
evaluate \eqref{eq:hydrophobicAttraction} using the identities
\begin{align}
  \label{eq:recipforcetorque}
  \FF_i^{\text{hydro}} = \int_{\Gamma_i} \JJ_i \,\dif s,\quad
  T_i^{\text{hydro}}    = \int_{\Gamma_i} 
    (\xx - \aa_i)^{\perp} \cdot \JJ_i  \,\dif s,
\end{align}
where
\begin{align}
  \label{eq:jumpstress1}
  \JJ_{i} = 2\gamma\rho^{-1} \sigma v_i \nnu + 
    2\gamma\rho \frac{d\sigma}{ds} \frac{dv_i}{ds} \nnu -
    2\gamma\rho \frac{d\sigma}{ds} \frac{dv_i}{d\nu} \ttau.
\end{align}
Here, $\ttau$ and $\frac{d}{ds}$ are the unit tangent  and arc length
derivative for $\Gamma_i$, respectively, and
\begin{equation}
\label{eq:uidef}
\begin{aligned}
    u(\xx) &= u_i(\xx) + v_i(\xx),\\
    u_i(\xx) &= \frac{1}{2\pi} \int_{\Gamma_i} \pderiv{}{\nnu_\yy}
    K_0 \left(\frac{|\xx - \yy|}{\rho}\right) \sigma(\yy) \, \dif s_\yy,
    \quad \xx \in \mathbb{R}^2.
\end{aligned}
\end{equation}
That is, $v_i(\xx)$ is the double-layer potential~\eqref{eq:HAP_DLP} with 
$\Gamma_i$ excluded from $\bd\Omega$. The advantage in
using \eqref{eq:recipforcetorque} over using \eqref{eq:hydrophobicAttraction} is that the components
$\sigma$, $v_i$, $\frac{d\sigma}{ds}$, and $\frac{dv_i}{ds}$ of $\JJ_i$ are smooth functions,
whereas the components of \eqref{eq:stress} are singular integrals.

@@

Let $U_i$ be the interior of the particle indexed by $i$. For $\xx_0
\in \Gamma_i$ and an arbitrary function $g(\xx)$, the notation
\begin{align}
  \jump{g}(\xx_0) = \lim_{\substack{\xx \to \xx_0 \\ \xx \in U_i^c}}g(\xx)  - 
                    \lim_{\substack{\xx \to \xx_0 \\ \xx \in U_i}}g(\xx),
\end{align}
denotes the jump of the limits of $g(\xx)$ taken from the outside to
the inside of $\Gamma_i$.
%\begin{align}
%\lim_{\xx \to \xx_0^+}g(\xx),\quad 
%\lim_{\xx \to \xx_0^-}g(\xx), \quad
%  \jump{g}(\xx_0) = \lim_{\xx \to \xx_0^+}g(\xx)  - \lim_{\xx \to \xx_0^-}g(\xx)
%\end{align}
%denotes the limit taken from points in $U_i^c$,  the limit taken from
%points in $U_i$, and the jump, respectively.  
\begin{lemma}
\begin{align}
  \label{eq:prejump}
  \FF_i = \int_{\Gamma_i} \jump{{\bf T}_2  \nnu}  \, \dif s,\quad
  T_i = \int_{\Gamma_i} (\xx - \aa_i)^{\perp} \cdot \jump{{\bf T}_2 \nnu} \, \dif s.
\end{align}
\end{lemma}
\begin{proof}
To show~\eqref{eq:prejump}, we expand
\begin{align}
  \FF_i = \int_{\Gamma_i} {\bf T}_1\nnu + 
    {\bf T}_2\nnu + {\bf T}_3\nnu\,\dif s.
\end{align}
By~\eqref{eq:uidef} and~\eqref{eq:decompdivfree}, we have that $u_i$ is
smooth and $\nabla \cdot {\bf T}_1 = \mathbf{0}$ in $\mathbb{R}^n
\setminus U_i$. Similarly, $v_i = u - u_i$ is smooth and $\nabla \cdot
{\bf T}_3 = \mathbf{0}$ in $U_i$. By the divergence theorem,  
\begin{align}
  \int_{\Gamma_i}  {\bf T}_1\nnu \,\dif s
  = -\int_{\mathbb{R}^n \setminus U_i} \nabla \cdot {\bf T}_1 \,\dif \xx = \mathbf{0},\quad
    \int_{\Gamma_i}  {\bf T}_3 \nnu\,\dif s
  = \int_{U_i} \nabla \cdot {\bf T}_3 \,\dif \xx = \mathbf{0}.
\end{align}
Finally, $u_i$ and $v_i$ are smooth and $\nabla \cdot {\bf T}_2= 0$ in $U_i$. This gives
\begin{align}
  \mathbf{0} = \int_{U_i} \nabla \cdot {\bf T}_2 \, \dif \xx = -\int_{\Gamma_i}  
    ({\bf T}_2 \nnu)^-\,\dif s,
\end{align}
where the superscript denotes the limit taken from in $U_i$.  
Combining the previous four equations gives the first equation in \eqref{eq:prejump}.
The argument for the second equation in \eqref{eq:prejump} is identical. 
\end{proof}

We have the jump relations for \eqref{eq:uidef}: 
\begin{equation}
\label{eq:DLjump}
\jump{u_i} = \sigma, \quad
\jump{v_i} = 0, \quad
\jump{\nabla u_i} = \frac{\dif \sigma}{\dif s}\ttau,\quad
\jump{\nabla v_i} = \mathbf{0},
\end{equation}
on $\Gamma_i$. We get 
\begin{align*}
  \jump{{\bf T}_2\nnu}   &= \jump{\mathbf{S}(u_i,v_i)\nnu  +\mathbf{S}(v_i,u_i)\nnu} \\
  &= \jump{( 2\gamma\rho^{-1} u_i v_i \mathbf{I} + 2\gamma\rho \nabla u_i \cdot \nabla v_i \mathbf{I} 
- 2\gamma\rho \nabla u_i  \nabla v_i^T - 2\gamma\rho \nabla v_i \nabla u_i^T)  \nnu}\\
&= 2\gamma\rho^{-1} \sigma v_i \nnu + 2\gamma\rho \frac{d\sigma }{ds}\frac{dv_i }{ds} \nnu
- 2\gamma\rho \ \frac{d\sigma }{ds} \frac{dv_i }{d\nu} \ttau = \JJ_i,
\end{align*}
Combing  this with \eqref{eq:prejump}  gives  \eqref{eq:recipforcetorque}
as required.

@@

\begin{table}
\caption{Friction Coefficients}
\centering    
\begin{tabular}{c c c c c c c c }
 $\chi$ & 0.0020   &  0.0025 &  0.0030 &  0.0035 &  0.0040 & 0.0045 & 0.0050  \\
\hline                    
$\dot\gamma$ (ns$^{-1}$)        & 0.0033 & 0.0041 & 0.0049 & 0.0057 & 0.0066 & 0.0074 & 0.0082\\
%$F_{f}$ (pN nm$^{-1}$)           & 1.84 & 4.90 & 11.36 & 8.16 & 6.14 &  17.69 &  24.33 \\[1ex]
%$\dot\theta$ (nm ns$^{-1}$)           & $6.84\times10^{-5}$ & $1.31\times10^{-4}$ & $3.46\times10^{-4}$ & $2.32\times10^{-4}$ & $1.81\times10^{-4}$ & $4.91\times10^{-4}$ & $5.83\times10^{-4}$\\ [1ex]
$b$ (pN$\cdot$ns nm$^{-2}$)    & 3.56 & 3.53 & 3.12 & 3.61 & 3.57 & 3.56 & 3.58\\ 
\hline    
\end{tabular} 
\label{table1}
\end{table}
